---
title: "Lab4"
author: "Toko Michioka"
date: "2024-09-26"
output: "github_document"
---

```{r}
library(car)
library(dplyr)
library(tidyr)
library(ggplot2)
library(psych)
library(bruceR)

```

# Load dataset
```{r}

lab4data <- read.csv("/Users/toko/Documents/GitHub/Lab4/skewed.csv")


```


# Normality
## Normality Plots
```{r}


ggplot(lab4data, aes(x = Performance)) + geom_histogram(binwidth = 20) + theme_classic()

ggplot(lab4data, aes(x = Performance)) + geom_density(adjust = 2)  + theme_classic()

qq<-ggplot(lab4data, aes(sample = Performance)) + geom_qq()  + theme_classic()

qq+ geom_qq_line()


#functions below are base R functions, they can do the job, but not as customizable as ggplot

hist(lab4data$Performance)

qqnorm(lab4data$Performance, col = "steelblue", lwd = 2)

plot(density(lab4data$Performance, na.rm = TRUE, bw = 90),  lwd=2, main = "")

```

### Normality Plots by Group
```{r}

#a fun function to plot the violin plot by group, but it's not part of ggplot
violinBy(Performance ~ Group, data = lab4data, rain= TRUE, vertical = FALSE)


#Find a way to plot the histograms, density, and qq plots by groups using ggplot
ggplot(lab4data, aes(x = Performance)) + geom_histogram(binwidth = 20) + facet_wrap (~Group) + theme_classic()

ggplot(lab4data, aes(x = Performance)) + geom_density(adjust = 2)  + facet_wrap (~Group) + theme_classic()

qq<-ggplot(lab4data, aes(sample = Performance)) + geom_qq()  + theme_classic()

qq+ geom_qq_line()

#Use ggplot to create a violin plot by groups


```


## Normality Tests
```{r}

describe(lab4data$Performance)

shapiro.test(lab4data$Performance)



```

### Normality Tests by Group
```{r}

#Use the describeBy() function to get skewness and kurtosis by group

?describeBy() 
describeBy(Performance ~ Group, data = lab4data)

#Use the group by function to get shapiro test results by group
lab4data %>%
  group_by(Group) %>%
  summarize(W = shapiro.test(Performance)$statistic, p_value = shapiro.test(Performance)$p.value)

#Use the filter function to get both





```
# Q1: What is your overall conclusion about this variable's normality? Why?
This variable's normality is met since  all of the p-values are over the threshold = .05, meaning it's insignificant.


# Equal Variance between Groups
## Descrptive Variance
```{r}
lab4data_clean<-drop_na(lab4data)
var(lab4data$Performance)

lab4data_clean %>%
  group_by(Group) %>%
  summarize(variance = var(Performance))


```

## Equal Variance Test
```{r}


leveneTest(Performance~Group, lab4data)

MANOVA(lab4data, dv = "Performance", between = "Group")

#What if you want to test equal variance between 2 groups specifically? 




```

# Q2: Overall, does it meet the equal variance assumption across the groups? Why?
No, because the p-value (.023) is under the threshold (.05), meaning it is insignificant. Therefore, it does not meet the equal variance assumption across the groups.

# Transformation
```{r}

#if any of the assumption is not met, use transformation 

lab4data$Performance_log <- log10(lab4data$Performance)
```

```{r}


ggplot(lab4data, aes(x = Performance_log)) + geom_histogram(binwidth = 0.25) + facet_wrap (~Group) + theme_classic()

ggplot(lab4data, aes(x = Performance_log)) + geom_density(adjust = 2)  + facet_wrap (~Group) + theme_classic()

qq<-ggplot(lab4data, aes(sample = Performance_log)) + geom_qq()  + theme_classic()

qq+ geom_qq_line()

describe(lab4data$Performance_log)

shapiro.test(lab4data$Performance_log)

describeBy(Performance_log ~ Group, data = lab4data)

lab4data %>%
  group_by(Group) %>%
  summarize(W = shapiro.test(Performance_log)$statistic, p_value = shapiro.test(Performance_log)$p.value)

lab4data_clean<-drop_na(lab4data)
var(lab4data$Performance_log)

lab4data_clean %>%
  group_by(Group) %>%
  summarize(variance = var(Performance_log))

leveneTest(Performance_log~Group, lab4data)

MANOVA(lab4data, dv = "Performance_log", between = "Group")

lab4dataConG1<-lab4data %>%
  filter(Group == "Control" | Group == "G1")

leveneTest(Performance_log~Group, lab4dataConG1)
```

# Q3: Run the above tests again with the transformed outcome. Compare the differences in results. 
Before we made the transformation, the assumption of normality was violated. After we transformed the data logarithmically and corrected bandwidth, the data appeared normal.



