---
title: "Lab4"
author: "Your Name"
date: "2024-09-26"
output: "github_document"
---

```{r}
library(car)
library(dplyr)
library(tidyr)
library(ggplot2)
library(psych)
library(bruceR)

```

# Load dataset
```{r}

lab4data <- read.csv("C:/Users/Colin/Downloads/skewed.csv")


```


# Normality
## Normality Plots
```{r}


ggplot(lab4data, aes(x = Performance)) + geom_histogram(binwidth = 20) + theme_classic()

ggplot(lab4data, aes(x = Performance)) + geom_density(adjust = 2)  + theme_classic()

qq<-ggplot(lab4data, aes(sample = Performance)) + geom_qq()  + theme_classic()

qq+ geom_qq_line()




qqnorm(lab4data$Performance, col = "steelblue", lwd = 2)

plot(density(lab4data$Performance, na.rm = TRUE, bw = 90),  lwd=2, main = "")

```

### Normality Plots by Group
```{r}

#a fun function to plot the violin plot by group, but it's not part of ggplot
#violinBy(Performance ~ Group, data = lab4data, rain= TRUE, vertical = FALSE)


#Find a way to plot the histograms, density, and qq plots by groups using ggplot


#Use ggplot to create a violin plot by groups


```


## Normality Tests
```{r}

describe(lab4data$Performance)

shapiro.test(lab4data$Performance)



```

### Normality Tests by Group
```{r}

#Use the describeBy() function to get skewness and kurtosis by group


#Use the group by function to get shapiro test results by group
lab4data %>%
  group_by(Group) %>%
  summarize(W = shapiro.test(Performance)$statistic)

#Use the filter function to get both





```
# Q1: What is your overall conclusion about this variable's normality? Why?



# Equal Variance between Groups
## Descrptive Variance
```{r}

var(lab4data$Performance)

lab4data %>%
  group_by(Group) %>%
  summarize(...)


```
## Equal Variance Test
```{r}


leveneTest(Performance~Group, lab4data)

MANOVA(lab4data, dv = "Performance", between = "Group")

#What if you want to test equal variance between 2 groups specifically? 

lab4dataConG1<-lab4data %>%
  filter(Group == "Control" | Group == "G1")

leveneTest(Performance~Group, lab4dataConG1)


```

# Q2: Overall, does it meet the equal variance assumption across the groups? Why?

# Transformation
```{r}

#if any of the assumption is not met, use transformation 

lab4data$Performance_log <- log10(lab4data$Performance)



```

# Q3: Run the above tests again with the transformed outcome. Compare the differences in results. 